---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

#############
Exercise 3-1
#############

We have 2 continous variable from hicker smarthphone.
dist -> Distance (in km)
alti -> Altitude (in m)

a) Calculate the arithmetic mean and median for both distance and altitude

Mean, Median for Distance

```{r}
dist<-c(12.5,29.9,14.8,18.7,7.6,16.2,16.5,27.4,12.1,17.5)
mean(dist)
median(dist)
```

Mean, Median for Altitude 

```{r}
alti<-c(342,1245,502,555,398,670,796,912,238,466)
mean(alti)
median(alti)
```

b) Determine the first and third quartiles for both distance and the altitude variables
Discuss the shape of the distribution give the result of (a) and (b)

Quartiles for distance

```{r}
quantile(dist,probs=c(0.25,0.75),type=2)
```
To achieve correct quantile type 2 algorithm was used (default is 7).


Quartiles for distance

```{r}
quantile(alti,probs=c(0.25,0.75),type=2)
```

What does it means? 
Distance. Mean is higher than median and mean in higher range of 25% and 75% quantile. It means 
there were some trip(s) with longer length than others.

Altitude. Mean is in the middle of 25% and 75% range. Mean is much higher than average, so we have some trips with mich higher 
altitude move then others.


Interquartile range for distance
```{r}
  quantile(dist,probs=c(0.75),type=2) - quantile(dist,probs=c(0.25),type=2)
```
Interquartile range for distance
```{r}
  quantile(alti,probs=c(0.75),type=2) - quantile(alti,probs=c(0.25),type=2)
```


Absolute median deviation

1. For each observation do abs(x-mean). Output in vector t.
2. Sum all values in t.
3. Divide by len(t)

Distance.
```{r}
dist<-c(12.5,29.9,14.8,18.7,7.6,16.2,16.5,27.4,12.1,17.5)
m<-median(dist)
dist = dist - m
dist = abs(dist)
sum(dist)/length(dist)
```
e) Boxplots

Boxplot for distance 
```{r}
boxplot(dist,range=1.5)
```

Boxplot for altitude
```{r}
boxplot(alti,range=1.5)
```

f) Parameters for grouped data. Calculate the weighted arithmetic data, weighted median

Weight arithmetic mean 
```{r}
weighted.mean(c(10,17.5,25),c(0.4,0.4,0.2))
```

Weight median (values are equally distributed)
Sum all relative frequencies and stop when > 0.5
median = 15 + ((0.5) - 0.4 /0.4) * 5 = 15 + (0.1/0.4) * 5 = 15 + 0.25 * 5 = 15 + 1.25 = 16.25 

Is there any way in R to support this automatically ?





